---
title: "covid19 visulization"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load covid-19 data and libraries


```{r pressure, echo=FALSE}
library(ggplot2)
library(ggplot2)
#install.packages("anytime")
library(anytime)
#install.packages("usmap")
library(usmap)
library(tidyverse)
library(dplyr)
theme_set(theme_bw())
```






# Data preprocessing
```{r pressure, echo=FALSE}
df <- read_csv("covid_19_data.csv",
col_types = cols(ObservationDate = col_date(format = "%m/%d/%Y")))

# rename the column name
colnames(df) <- c("SNo","Date","Province","Country","Update","Confirmed","Deaths","Recovered")

df <- df %>%
    mutate(Province = case_when(
        Province == "France"~"",
        Province == "United Kingdom"~"",
        Province == "UK"~"",
        Province == "Netherlands"~"",
        Province == "Denmark"~"",
        Province == "US"~"",
        Province == "Canada"~"",
        Province == "Russia"~"",
        Province == "Italy"~"",
        TRUE ~ Province))

#check all missing values
sapply(df, function(k) sum(is.na(k)))

# drop missing values
df1<-na.omit(df)
df1
```




# Exploratory Data Analysis of Covid-19 confirmed cases

# picked  6 countries from different continents to examine the daily death 

# US daily death from 2020-2021

```{r pressure, echo=FALSE}

us <- df1 %>%
    filter(Country == "US") %>%
    group_by(Date) %>%
    summarize(total_deaths = sum(Deaths)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1))


ggplot(subset(us,Date >= anydate("03/15/2020") & Date < anydate("04/14/2021")), aes(x = Date, y = daily_deaths, color= daily_deaths)) +
    geom_col() +
    ylab(label = "Daily Deaths") + 
    xlab("Date")

```


# Canada daily death from 2020-2021


```{r pressure, echo=FALSE}

CAN <- df1 %>%
    filter(Country == "Canada") %>%
    group_by(Date) %>%
    summarize(total_deaths = sum(Deaths)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1))


ggplot(subset(CAN,Date >= anydate("03/15/2020") & Date < anydate("04/14/2021")), aes(x = Date, y = daily_deaths, color=daily_deaths )) +
    geom_col() +
    ylab(label = "Daily Deaths") + 
    xlab("Date")

```


# Brazil daily death from 2020-2021

```{r pressure, echo=FALSE}

bra <- df1 %>%
    filter(Country == "Brazil") %>%
    group_by(Date) %>%
    summarize(total_deaths = sum(Deaths)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1))


ggplot(subset(bra,Date >= anydate("02/15/2020") & Date < anydate("07/14/2021")), aes(x = Date, y = daily_deaths, color=daily_deaths )) +
    geom_col() +
    ylab(label = "Daily Deaths") + 
    xlab("Date")

```
# China daily death from 2020-2021

```{r pressure, echo=FALSE}

china <- df1 %>%
    filter(Country == "Mainland China") %>%
    group_by(Date) %>%
    summarize(total_deaths = sum(Deaths)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1))


ggplot(subset(china,Date >= anydate("01/15/2020") & Date < anydate("03/14/2021")), aes(x = Date, y = daily_deaths, color=daily_deaths )) +
    geom_col() +
    ylab(label = "Daily Deaths") + 
    xlab("Date")

```
## upclose look into china highest spike

```{r pressure, echo=FALSE}

china <- df1 %>%
    filter(Country == "Mainland China") %>%
    group_by(Date) %>%
    summarize(total_deaths = sum(Deaths)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1))


ggplot(subset(china,Date >= anydate("04/01/2020") & Date < anydate("05/01/2020")), aes(x = Date, y = daily_deaths, color=daily_deaths )) +
    geom_col() +
    ylab(label = "Daily Deaths") + 
    xlab("Date")

```



# Australia

```{r pressure, echo=FALSE}

aus <- df1 %>%
    filter(Country == "Australia") %>%
    group_by(Date) %>%
    summarize(total_deaths = sum(Deaths)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1))


ggplot(subset(aus,Date >= anydate("03/15/2020") & Date < anydate("04/14/2021")), aes(x = Date, y = daily_deaths, color=daily_deaths )) +
    geom_col() +
    ylab(label = "Daily Deaths") + 
    xlab("Date")

```

```{r pressure, echo=FALSE}

mex <- df1 %>%
    filter(Country == "Mexico") %>%
    group_by(Date) %>%
    summarize(total_deaths = sum(Deaths)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1))


ggplot(subset(mex,Date >= anydate("03/15/2020") & Date < anydate("04/14/2021")), aes(x = Date, y = daily_deaths, color=daily_deaths )) +
    geom_col() +
    ylab(label = "Daily Deaths") + 
    xlab("Date")

```



# Top total confirmed cases  across the globe

```{r}
df1_country <- df1 %>%
    group_by(Country, Date) %>%
    summarize(total_deaths = sum(Deaths), total_confirmed = sum(Confirmed)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1), 
           daily_confirmed = total_confirmed - lag(total_confirmed,1))

last <- df1_country %>%
    arrange(desc(Date)) %>%
    filter(!duplicated(Country)) %>%
    arrange(desc(total_confirmed))

print(last[1:20,])
last$Country[1:8]

ggplot(filter(df1_country, Country %in% last$Country[1:10], Date< anydate("06/15/2020")),
      aes(x = Date, y = total_confirmed, color = Country)) + geom_line()

```

# Top total death cases  across the globe early stage and later stage of the pandemic


```{r}
df1_country <- df1 %>%
    group_by(Country, Date) %>%
    summarize(total_deaths = sum(Deaths), total_confirmed = sum(Confirmed)) %>%
    mutate(weekday = weekdays(Date), daily_deaths = total_deaths - lag(total_deaths,1), 
           daily_confirmed = total_confirmed - lag(total_confirmed,1))

last <- df1_country %>%
    arrange(desc(Date)) %>%
    filter(!duplicated(Country)) %>%
    arrange(desc(total_confirmed))

print(last[1:20,])
last$Country[1:8]

g1<-ggplot(filter(df1_country, Country %in% last$Country[1:10], Date < anydate("05/15/2020")),
      aes(x = Date, y = total_deaths, color = Country)) + geom_line()

g2<-ggplot(filter(df1_country, Country %in% last$Country[1:10], Date > anydate("04/15/2020")),
      aes(x = Date, y = total_deaths, color = Country)) + geom_line()

library(gridExtra)
grid.arrange(g1,g2)

```



```{r}
# World confirmed with counts in log10 scale
ggplot(df1, aes(x=Date, y=Confirmed,color= Confirmed)) + geom_line() +
  theme_classic() +
  labs(title = "Covid-19 Global Confirmed Cases in log scale", x= "Days", y= "Daily confirmed cases  (log scale)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(trans="log10")

```
# Mapping Covid-19 confirmed cases

```{r}
# Map
library(tmap)
data(World)
class(World)
countrytotal <- df1 %>% group_by(Country) %>% summarize(cumulativeconfirmed=sum(Confirmed), cumulativedeaths=sum(Deaths), cumulativerecovered=sum(Recovered))

# Combine basemap data to covid data
countrytotal$Country[!countrytotal$Country %in% World$name]
list <- which(!countrytotal$Country %in% World$name)
countrytotal$country <- as.character(countrytotal$Country)


countrytotal$Country[!countrytotal$Country %in% World$name]
World$country <- World$name
worldmap <- left_join(World, countrytotal, by="country")
worldmap$cumulativeconfirmed[is.na(worldmap$cumulativeconfirmed)] <- 0



# Map
ggplot(data = worldmap) + geom_sf(aes(fill=cumulativeconfirmed), color="black") +
  ggtitle("World Map of Confirmed Covid Cases",
          subtitle="Total Cases on May 29, 2021") +
  theme_bw()




```



